---
title: "Air pollution comparison with respect to traffic volume and milage covered by vehicles"
author: "khalifa alshamsi"
date: "10/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

july 2020 dataset is used by US department of fedral Highway administration \url{https://www.fhwa.dot.gov/policyinformation/travel_monitoring/tvt.cfm}

## Traffic Volume Trends 2020

```{r}

library(tidyverse)

tv <- read.csv("tv.csv")

ggplot(data = tv, aes(x = year_record, y = tmonth)) +
  geom_line()+
    labs(title = "Travel volume for the current month (1995-2020)",
         x = "Year of observation",
         y = "Number of trafic vehicles")

ggplot(data = tv, aes(x = year_record, y = yearToDate)) +
  geom_line()+
    labs(title = "Travel volume for the cumulative yearly total (1995-2020)",
         x = "Year of observation",
         y = "Number of trafic vehicles")

ggplot(data = tv, aes(x = year_record, y = moving)) +
  geom_line()+
    labs(title = "Travel volume for the cumulative yearly total (1995-2020)",
         x = "Year of observation",
         y = "Number of trafic vehicles")

```

According to travel volume for the current month in the year 2019 the number of vehicles increased a lot in 2020 and as the number of vehicles increases the air pollution also increases. 

with the 2nd graph with respect to yearly trafic observations we can interpret that fom continous increase in pollution in the year 2020 the level of traffic vehicles decreases as in the year 2000. 

The moving vehicles in july 2020 is decreased and equal to 2014.

This total includes  86.3 billion vehicle-miles on rural roads and  176.2 billion vehicle-miles on urban roads and streets.
Cumulative Travel changed by  -15.7% (-297.2 billion vehicle miles).


## Rural Highways- Average Daily Vehicle-Distance Traveled

```{r}

rv <- read.csv("rv.csv")

ggplot(data = rv, aes(x = month, y = X2020)) +
  geom_boxplot()+
    labs(title = "Rural Highways- Average Daily Vehicle-Distance Traveled(Billion Miles) in 2020",
         x = "Month",
         y = "Average Daily Vehicle-Distance Traveled(Billion Miles)")

ggplot(data = rv, aes(x = month, y = X2019)) +
  geom_boxplot()+
    labs(title = "Rural Highways- Average Daily Vehicle-Distance Traveled(Billion Miles) in 2019",
         x = "Month",
         y = "Average Daily Vehicle-Distance Traveled(Billion Miles)")

```

In 2020 the air vehicle distance travelled is also decreased with respect to 2019 and the rural travel index is about 2.5 for july for the year of 2020 and for 2019 it it is approx 3. which shows that this year the distance travelled by vehicles is low which shows that the air pollution decreases.

## Urban Highways- Average Daily Vehicle-Distance Traveled

```{r}

uv <- read.csv("uv.csv")

ggplot(data = rv, aes(x = month, y = X2020)) +
  geom_point()+
    labs(title = "Urban Highways- Average Daily Vehicle-Distance Traveled(Billion Miles) in 2020",
         x = "Month",
         y = "Average Daily Vehicle-Distance Traveled(Billion Miles)")

ggplot(data = rv, aes(x = month, y = X2019)) +
  geom_point()+
    labs(title = "Urban Highways- Average Daily Vehicle-Distance Traveled(Billion Miles) in 2019",
         x = "Month",
         y = "Average Daily Vehicle-Distance Traveled(Billion Miles)")

```


Traffic Volume Trends is a monthly report based on hourly traffic count data. These data, collected at approximately 
5,000 continuous traffic counting locations nationwide, are used to determine the percent change in traffic for the 
current month compared to the same month in the previous year. This percent change is applied to the travel for the 
same month of the previous year to obtain an estimate of travel for the current month.Because of the limited sample 
sizes, caution should be used with these estimates.  The Highway Performance Monitoring System provides more 
accurate information on an annual basis.

## Seasonally Adjusted Vehicle Miles Traveled by Month

Seasonally-adjusted VMT data enables comparisons with any other month in any other year. Analysis of seasonally-adjusted VMT is an alternative to analysis of unadjusted VMT, which traditionally uses comparisons of a month to the same month in previous years to determine trends. Seasonally adjusted VMT data enables not only same monthly travel pattern comparisons from different years, but comparisons with any other month.


```{r}

vtm <- read.csv("vtm.csv")
head(vtm)
names(vtm)

ggplot(data = vtm, aes(x = Date, y = Seasonally.Adjusted.VMT..2000.to.present.)) +
  geom_point()+
    labs(title = "Seasonally Adjusted Vehicle Miles Traveled by Month",
         x = "Month",
         y = "VTM")+theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks = element_blank())

```

Discussion

Travel on all roads and streets changed by -11.2% (-33.2
billion vehicle miles) for July 2020 as compared with July
2019. Travel for the month is estimated to be 262.4 billion
vehicle miles.
The seasonally adjusted vehicle miles traveled for July 2020
is 239.7 billion miles, a -12.5% (-34.4 billion vehicle miles)
decline from July 2019. It also represents 4% increase (9.2
billion vehicle miles) compared with June 2020.
Cumulative Travel for 2020 changed by -15.7% (-297.2
billion vehicle miles). The Cumulative estimate for the year
is 1,593.2 billion vehicle miles of travel

hh


